\section{Eigenvalues and Eigenvectors}

\subsection{Eigenvalues and Eigenvectors}
\paragraph{Definitions}
Suppose \(T \in L(V,V):\)
\begin{enumerate}
    \item if \(T(\vv) = \alpha\vv, \vv \in V, \alpha \in \bF, \vv \neq 0\) then \(\alpha\) is an eigenvalue of \(T, \vv\) is an eigenvector for \(T\) corresponding to \(\alpha\).
    \item If \(\lambda\) is an eigenvalue of \(T\), then the eigenspace of \(T, E_\lambda(T) = \{\vv \in V: T(\vv) = \lambda\vv \}\).
    \item We call the set of all eigenvalues of \(T\) the spectrum of \(T\).
\end{enumerate}

\paragraph{Notes}
\begin{enumerate}
    \item Eigenvectors are never zero.
    \item \(E_\lambda(T) = \{\vzero\} \cup \{\text{ all eigenvectors corresponding to } \lambda\}.\)
    \item Each eigenspace is invariant under T.
\end{enumerate}

\paragraph{Properties of Eigenvalues}
Let \(T: V \to V\) be linear. Then
\begin{enumerate}
    \item \(E_\lambda(T) = \ker(\lambda\id - T)\).
    \item If \(\lambda_1, \dots, \lambda_k\) are distinct eigenvalues and \(T(\vv_i) = \lambda_i\vv_i\), then \(\vv_1, \dots, \vv_k\) are linearly independent.
    \item If \(\lambda \neq \mu\) then \(E_\lambda(T) \cap E_\mu(T) = \{\vzero\}\).
\end{enumerate}

\paragraph{Diagonalisable}
A matrix \(A \in M_{p,p}(\bF)\) is diagonalisable over \(\bF\) if there is an invertible matrix \(P \in \GL(p,\bF)\) such that \(P^{-1}AP\) is diagonal - that is, A is similar over \(\bF\) to a diagonal matrix. A linear map \(T: V \to V\) is diagonalisable if there is a basis of \(V\) with respect to which the matrix of \(T\) is diagonal.

\paragraph{Diagonalisability Regarding Basis}
If \(T \in L(V,v)\) where \(V\) is a finite dimensional vector space over \(\bF\), then \(T\) is diagonalisable if and only if \(V\) has a basis whose elements are all eigenvectors of \(T\). Similarly, \(A \in M_{p,p}(\bF)\) is diagonalisable if and only if \(\bF^p\) has a basis consisting of eigenvectors of \(A\). 

\paragraph{Diagonalisability Regarding Eigenvalues}
A \(p \times p\) matrix with \(p\) distinct eigenvalues is diagonalisable.

\subsection{The Characteristic Polynomial}
\paragraph{Definition}
If \(A\) is a \(p \times p\) matrix over \(\bF\) then the characteristic polynomial of \(A\) is
\[\charp_A(t) = \det(tI - A).\]

\paragraph{Similarity Invariant}
The characteristic polynomial is a similarity invariant.

\paragraph{Characteristic Polynomials of Maps}
If \(V\) is finite dimensional and \(T \in L(V,V)\) the characteristic polynomial of \(T, \charp_T(t)\), is defined to be \(\charp_A(t)\) for any matrix \(A\) of \(T\).

\paragraph{Properties with Linear Maps}
Suppose \(T \in L(V,V)\), with \(\dim V = n\). Then
\begin{enumerate}
    \item \(\lambda\) is an eigenvalue if and only if \(\charp_T(\lambda) = 0\).
    \item \(W \leq E_\lambda(T)\) implies \(W\) is \(T\)-invariant.
    \item \(\lambda\) is an eigenvalue if and only if \(\nullity(T - \lambda\id) > 0\).
\end{enumerate}

\paragraph{Eigenvalues with Basis}
Let \(T: V \to V\) be linear on finite dimensional space \(V\), and \(\basis\) be a basis for \(V\). Let \(A\) be the matrix of \(T\) with respect to \(\basis\). The following are equivalent:
\begin{itemize}
    \item \(\vv\) is an eigenvector of \(T\) corresponding to eigenvalue \(\lambda\);
    \item the coordinate vector \([\vv]_\basis\) is an eigenvector of \(A\) corresponding to eigenvalue \(\lambda\).
\end{itemize}

\paragraph{Diagonalisation Algorithm}
Given matrix \(A\)
\begin{enumerate}
    \item Calculate \(\charp_A(t)\).
    \item Find all roots \(\alpha\) of \(\charp_A(t)\).
    \item For each \(\alpha\), calculate \(\dim(E_\alpha(A))\).
    \item If \(\sum_\alpha \dim(E_\alpha(A)) = n\) then \(A\) is diagonalizable.
    \item If \(A\) is diagonalisable construct \(P\), whose columns are a basis of \(V\) consisting of eigenvectors of \(A\) so that \(D = P^{-1}AP\) is diagonal.
\end{enumerate}

\subsection{Multiplicities}
\paragraph{Geometric and Algebraic Multiplicity}
Suppose \(T\) is a linear map on finite dimensional vector space \(V\). Suppose further that \(\lambda\) is an eigenvalue of \(T\) so that \(t - \lambda\) is a factor of the characteristic polynomial of \(T\). 
\begin{itemize}
    \item The geometric multiplicity (g.m.) of \(\lambda\) is \(\dim(E_\lambda(T))\).
    \item The algebraic multiplicity (a.m.) of \(\lambda\) is its multiplicity as a factor of \(\charp_T(t)\).
\end{itemize}
Let \(T: V \to V\) be linear on finite dimensional space \(V\) and \(\lambda\) an eigenvalue of \(T\). Then
\[1 \leq \text{geometric multiplicity } \lambda \leq \text{algebraic multiplicity } \lambda.\]

\paragraph{Solutions, Determinant and Trace with Eigenvalues}
Suppose \(A \in M_{p,p}(\bC)\). Then \(A\) has \(p\) eigenvalues \(\alpha_1, \dots, \alpha_p\) counting algebraic multiplicities. Also
\[\det(A) = \prod_{i=1}^p \alpha_i \text{ and } \tr(A) = \sum_{i=1}^p \alpha_i.\]

\paragraph{Equivalent Properties}
Let \(T: V \to V\) be linear on finite dimensional space \(V\). The following four statements are equivalent:
\begin{enumerate}
    \item \(T\) is diagonalisable.
    \item There is a basis for \(V\) consisting of eigenvectors for \(T\).
    \item \(V = E_{\lambda_1}(T) \oplus E_{\lambda_2}(T) \oplus \cdots \oplus E_{\lambda_k}(T)\) where \(\lambda_1, \dots, \lambda_k\) are the distinct eigenvalues of \(T\).
    \item The sum of the geometric multiplicities of the distinct eigenvalues is \(\dim V\). That is,
    \[\sum_{j=1}^k \dim E_{\lambda_j}(T) = \dim V.\]
\end{enumerate}

\paragraph{Diagonalizable Distinct Roots}
Linear map \(T\) on a \(p\)-dimensional space \(V\) over \(\bF\) is diagonalizable if \(\charp_T(t)\) has \(p\) distinct roots in \(\bF\).